\textbf{Appendix}\\ Bound within Boundaries: How Well Do Protected Areas Match Movement Corridors of Their Most Mobile Protected Species?

### 2021-01-04 Chapter_0/04_Manuscript/Manuscript.tex
Implemented minor revisions

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell. Because our representation of water was dynamic we
collapsed all dynamic water maps into a single map using areas that were covered
by water in at least 10\% of the cases. We used the resulting map to also
calculate a layer returning the distance to water. Because the delta only covers
5\% of the KAZA-TFCA, we considered the use of a single water map to be
appropriate. To reduce the influence of outliers in predicted permeability
scores, we followed \cite{Squires.2013} and curtailed predicted scores between
the 1\textsuperscript{st} and 99\textsuperscript{th} percentile of their
original values. To compare permeability across different regions, we normalized
the permeability surface to a range between 0 (most impermeable) and 1 (most
permeable), and we determined median permeability within and outside the
KAZA-TFCA, within and outside formally protected areas, and within each of the
five KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify movement corridors of dispersing wild dogs, we specified source
points and calculated factorial least-cost paths (LCPs) as well as factorial
least-cost corridors (LCCs) among them \citep{Elliot.2014}. We generated source
points by overlaying the study area with a regular grid of points spaced at 100
km. We only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2}, which conforms with home-range requirements of African
wild dogs \citep{Pomilia.2015}. Finally, we defined centroids as source points
for those protected areas \(>\) 700 km\textsuperscript{2} that were not assigned
any source points from the regular grid. Because wild dogs residing outside of
protected areas are rare and unviable \citep{VanDerMeer.2014}, we consider our
selection of source points within protected areas to be appropriate. In total,
we generated 68 source points, which resulted in 2'278 unique pairwise
combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
LCPs and LCCs between source points using the R-package \textit{gdistance}
(further details in Appendix A.7). After computation, we tallied overlapping
LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
discrete paths, whereas LCCs return continuous corridors, we present both
methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

@@

\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)},
\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). Dispersers
avoided moving through water (\(\beta\) = -0.53, 95\% \(CI\) -0.79 to -0.27) but
selected for locations in its vicinity, although the latter effect was not
significant (\(\beta\) = -0.33, 95\% \(CI\) = -0.73 to 0.08). Dispersers avoided
areas that were densely covered by trees (\(\beta\) = -0.31, \(CI\) = -0.47 to
-0.15) and preferred areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\%
\(CI\) = 0.07 to 0.42). Finally, dispersers avoided areas that were influenced
by humans (\(\beta\) = -0.45, 95\% \(CI\) = -0.82 to -0.08). Except for
\textit{distance to water} (\(SD\)\textsubscript{RandomEffect} = 0.57), we
observed little variation between dispersal coalitions' relative selection
strengths (\(SD\)\textsubscript{RandomEffect} \(<\) 0.22 for all other
covariates, see also Figure S7).

@@

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (dark grey borders). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference, we also labeled some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

>>>>

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell. Because our representation of water was dynamic, we
collapsed all dynamic water maps into a single map using areas that were covered
by water in at least 10\% of the cases. We used the resulting map to also
calculate a layer returning the distance to water. Because the delta only covers
5\% of the KAZA-TFCA, we considered the use of a single water map to be
appropriate. To reduce the influence of outliers in predicted permeability
scores, we followed \cite{Squires.2013} and curtailed predicted scores between
the 1\textsuperscript{st} and 99\textsuperscript{th} percentile of their
original values. To compare permeability across different regions, we normalized
the permeability surface to a range between 0 (most impermeable) and 1 (most
permeable), and we determined median permeability within and outside the
KAZA-TFCA, within and outside formally protected areas, and within each of the
five KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify potential movement corridors of dispersing wild dogs, we specified
source points and calculated factorial least-cost paths (LCPs) as well as
factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To specify
source points, we applied the omnidirectional \textit{go-through} approach
proposed by \citep{Koen.2014} and placed 68 regularly spaced source points along
the map border. This approach has been shown to reduce potential biases arising
from the selection of inappropriate source locations \citep{Koen.2014,
Pitman.2017}. To assess the sensitivity of our results with respect to the
location of source points, we also applied an alternative method and placed 68
source points within protected areas that are large enough to sustain viable
wild dog populations (further details and corresponding results in Appendix
A.9). In either case, the 68 source points resulted in 2'278 unique pairwise
combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
LCPs and LCCs between source points using the R-package \textit{gdistance}
(details in Appendix A.6). After computation, we tallied overlapping
LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
discrete paths, whereas LCCs return continuous corridors, we present both
methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

@@

\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)},
\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). Dispersers
avoided moving through water (\(\beta\) = -0.53, 95\% \(CI\) -0.79 to -0.27) but
selected for locations in its vicinity, although the latter effect was not
significant (\(\beta\) = -0.33, 95\% \(CI\) = -0.73 to 0.08). Dispersers avoided
areas that were densely covered by trees (\(\beta\) = -0.31, \(CI\) = -0.47 to
-0.15) and preferred areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\%
\(CI\) = 0.07 to 0.42). Finally, dispersers avoided areas that were influenced
by humans (\(\beta\) = -0.45, 95\% \(CI\) = -0.82 to -0.08). Except for
\textit{distance to water} (\(SD\)\textsubscript{RandomEffect} = 0.57), we
observed little variation between dispersal coalitions' relative selection
strengths (\(SD\)\textsubscript{RandomEffect} \(<\) 0.22 for all other
covariates, see also Figure S8).

@@

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black semicircles along the map border) and
    corresponding least-cost paths between them. Continuous thin black lines
    indicate the borders of the KAZA-TFCA, whereas dashed black lines delineate
    country-borders. (b) Least-cost corridors between the same source points as
    illustrated in subfigure (a). For ease of spatial reference, we also labeled
    some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

### 2021-01-04 Chapter_0/04_Manuscript/Manuscript_Support.tex
Implemented minor revisions

%------------------------------------------------------------------------------
%  Appendix S6: Random Effects Variability
%------------------------------------------------------------------------------
\newpage
\section{Habitat Selection Model: Random Effects}
\begin{figure}[hbtp]
  \includegraphics[width = \textwidth]{99_RandomEffects}
  \caption{Plot of random effects showing variability across dispersal
  coalitions. The coalition ID refers to the individual in the dispersal
  coalition that was equipped with a GPS collar.}
  \label{RandomEffects}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S7: Least Cost Paths and Least Cost Corridors
%------------------------------------------------------------------------------
\newpage
\section{Identification of Least-Cost Paths \& Corridors}
\subsection{Least-Cost Paths}
We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
the (unscaled) permeability surface into a network of nodes to find shortest
effective distances between source points based on probabilities of moving from
cell to cell. In our case, the transition probability of moving between two
adjacent cells depended on their averaged permeability. We allowed individuals
to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
neighborhood) and applied a geographic correction to account for the fact that
diagonal neighbors were more remote than orthogonal neighbors. Because African
wild dogs have been observed to cover large dispersal distances
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
maximal effective cost. After computation, we tallied overlapping LCPs and
identified high-frequency routes.

@@

%  Appendix S8: AIC Model Selection

@@

>>>>

%------------------------------------------------------------------------------
%  Appendix S6: Least Cost Paths and Least Cost Corridors
%------------------------------------------------------------------------------
\newpage
\section{Identification of Least-Cost Paths \& Corridors}
\subsection{Least-Cost Paths}
We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
the (unscaled) permeability surface into a network of nodes to find shortest
effective distances between source points based on probabilities of moving from
cell to cell. In our case, the transition probability of moving between two
adjacent cells depended on their averaged permeability. We allowed individuals
to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
neighborhood) and applied a geographic correction to account for the fact that
diagonal neighbors were more remote than orthogonal neighbors. Because African
wild dogs have been observed to cover large dispersal distances
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
maximal effective cost. After computation, we tallied overlapping LCPs and
identified high-frequency routes.

@@

%  Appendix S7: AIC Model Selection

@@

%------------------------------------------------------------------------------
%  Appendix S8: Random Effects Variability
%------------------------------------------------------------------------------
\newpage
\section{Habitat Selection Model: Random Effects}
\begin{figure}[hbtp]
  \includegraphics[width = \textwidth]{99_RandomEffects}
  \caption{Plot of random effects showing variability across dispersal
  coalitions. The coalition ID refers to the individual in the dispersal
  coalition that was equipped with a GPS collar.}
  \label{RandomEffects}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S9: Alternative Source Points
%------------------------------------------------------------------------------
\newpage
\section{Alternative Source Points}
Because we were interested in assessing the sensitivity of our results with
respect to the location of source points, we implemented two competing
approaches to select source points. In a first run, we applied an
omnidirectional \textit{go-through} approach as proposed by \cite{Koen.2014} and
applied by \cite{Pitman.2017}. That is, we distributed 68 equally spaced source
points along the map border and we calculated least-cost paths and least-cost
corridors for all pairwise combinations of them. In contrast to other methods,
this approach has been shown to be less subjective and less sensitive to the
selection of unreasonable source points. Furthermore, such an approach enables
the identification of dispersal routes that do not necessarily connect patches
inhabited by the species. In a second run, we followed the approach described in
\cite{Elliot.2014} and overlayed the study area with a regular grid of points
spaced 100 km apart. From this regular grid we only considered those points that
fell within protected areas \(>\) 700 km\textsuperscript{2}, which conforms with
home-range requirements of African wild dogs \citep{Pomilia.2015}. Finally, we
defined centroids as source points for those protected areas \(>\) 700
km\textsuperscript{2} that were not assigned any source points from the regular
grid. Because wild dogs residing outside of protected areas are rare and
unviable \citep{VanDerMeer.2014}, we consider such a selection of source points
within protected areas to be reasonable. The second procedure also resulted in
68 source points, implying that we calculated 2'278 unique pairwise combinations
and therefore 2'278 unique LCPs and LCCs per approach.
\Cref{AlternativeSourcePoints} illustrates the resulting least-cost paths and
least-cost corridors and highlights that both methods produce qualitatively
similar results. In fact, map values of \Cref{AlternativeSourcePoints} (b1) and
(b2) correlate with \(r\) = 0.86. Nevertheless, one can see that connectivity
following the first approach gets somewhat  ``stretched'' towards the map
borders, which is expected given that the source points are all located at the
borders. Regardless of the method, we find that the major dispersal corridors
run within the borders of the KAZA-TFCA and that mainly impenetrable landscape
remains beyond its borders.

\begin{figure}[hbtp]
  \begin{center}
  \includegraphics[width = \textwidth]{99_LeastCost_Suppl}
  \caption{Comparison of least-cost paths (a) and least-cost corridors (b) using
  different source points. In subfigures (a1) and (b1), 68 source points were
  placed along the map border (black and white semicircles). In subfigures (a2)
  and (b2), on the other hand, source points (black and white circles) were
  placed in protected areas (gray) that were large enough to sustain viable wild
  dog populations (i.e. larger than 700 km\textsuperscript{2},
  \citep{Pomilia.2015}). Continuous thin black lines indicate the borders of the
  KAZA-TFCA, whereas dashed black lines delineate country-borders. For ease of
  spatial reference, we also labeled some national parks (NPs, in dark-grey).}
  \label{AlternativeSourcePoints}
  \end{center}
\end{figure}
